MPA Gap Analysis: IUCN and AquaMaps

### 2016-01-16 data_explore/mpa_analysis.Rmd
trying to get shapefiles saved and rasters burned... saving shapefile so we can use gdal_rasterize effectively

  writeRaster(rast_base, rast_base_file)
} else {
  rast_base <- raster(rast_base_file)
}

@@

``` {r set_up_spatially_marine_raster}
### import OHI regions polygons, to determine spatially-marine areas
dir_rgn   <- file.path(dir_N, 'git-annex/globalprep/spatial/v2015/data')
lyr_rgn <- 'regions_gcs'
poly_rgn <- readShapePoly(file.path(dir_rgn, lyr_rgn), 
                           proj4string = crs_wgs84)

### filter out all polygons that are land or landlocked, and flag the rest as spatially-marine = 1
poly_marine <- poly_rgn[!str_detect(poly_rgn@data$rgn_typ, 'land'), ]
poly_marine@data <- poly_marine@data %>%
  mutate(sp_mar_flag = 1)

# writePolyShape(poly_marine, file.path(dir_anx, 'explore_am_v_iucn/tmp_rasters/poly_sp_mar'))

# rast_sp_mar <- rasterize(poly_marine, rast_base, value = sp_mar_flag) ### pretty slow... consider gdal_rasterize?

# library(gdalUtils)
# rast_sp_mar <- gdal_rasterize(
#     src_datasource = file.path(dir_anx, 'explore_am_v_iucn/tmp_rasters/poly_sp_mar.shp')),
#     dst_filename = file.path(dir_anx, 'explore_am_v_iucn/tmp_rasters/poly_sp_mar.tif'), 
#       # destination for output
#     a = 'sp_mar_flag', 
#       # the attribute in the shapefile to be assigned to the cell values
#     te = c(-180, -90, 180, 90), 
#       # extents for output raster
#     tr = c(.01, .01),  
#       # resolution for output raster
#     tap = TRUE, 
#       # target aligned pixels - align coords of extent of output to values of -tr
#     a_nodata = NA, 
#       # nodata value for raster; otherwise they will be filled in with zeroes
#     output_Raster = TRUE, 
#       # return output as a RasterBrick? 
#     verbose = TRUE)

>>>>

  writeRaster(rast_base, rast_base_file, overwrite = TRUE)
} else {
  rast_base <- raster(rast_base_file)
}

@@

``` {r set_up_spatially_marine_raster}
### This raster (at 0.01 degree resolution) contains flags on whether a cell
### is "spatially marine" - i.e. overlaps (center?) with an ocean area.
### This is separate from the WDPA "marine" flag.

rast_marine_file <- file.path(dir_git, 'data/mpa_analysis/rast_marine.tif')
reload <- FALSE
if(reload == TRUE) {
### import OHI regions polygons, to determine spatially-marine areas
  poly_marine_file <- file.path(dir_anx, 'explore_am_v_iucn/tmp_rasters/poly_marine.shp')
  if(!file.exists(poly_marine_file)) {
    dir_rgn   <- file.path(dir_N, 'git-annex/globalprep/spatial/v2015/data')
    lyr_rgn <- 'regions_gcs'
    poly_rgn <- readShapePoly(file.path(dir_rgn, lyr_rgn), 
                               proj4string = crs_wgs84)

    ### filter out all polygons that are land or landlocked, and flag 
    ### the rest as spatially-marine = 1
    poly_marine <- poly_rgn[!str_detect(poly_rgn@data$rgn_typ, 'land'), ]
    poly_marine@data <- poly_marine@data %>%
      mutate(sp_mar = 1)

    # writePolyShape(poly_marine, poly_marine_file)
    writeOGR(obj    = poly_marine,
             dsn    = path.expand(dirname(poly_marine_file)),
             layer  = str_replace(basename(poly_marine_file), '.shp', ''),
             driver = 'ESRI Shapefile',
             overwrite_layer = TRUE,
             verbose = TRUE)

    } else {
    poly_marine <- readShapePoly(poly_marine_file, 
                               proj4string = crs_wgs84)
  }

  # rast_marine <- rasterize(poly_marine, rast_base, value = sp_mar_flag) 
  ### excruciatingly slow (interminable?)... consider gdal_rasterize?

  library(gdalUtils)
  rast_marine <- gdal_rasterize(
      src_datasource = poly_marine_file),
      dst_filename   = rast_marine_file), 
        # destination for output
      a = 'sp_mar_flag', 
        # the attribute in the shapefile to be assigned to the cell values
      te = c(-180, -90, 180, 90), 
        # extents for output raster
      tr = c(.01, .01),  
        # resolution for output raster
      tap = TRUE, 
        # target aligned pixels - align coords of extent of output to values of -tr
      a_nodata = NA, 
        # nodata value for raster; otherwise they will be filled in with zeroes
      output_Raster = TRUE, 
        # return output as a RasterBrick? 
      verbose = TRUE)
}
